---
title: "Models of Neighborhood Change in Salt Lake City, Utah"
output: 
  flexdashboard::flex_dashboard:
    theme: bootstrap
    source: embed
    smart: false
runtime: shiny
---

    

```{r global, echo=F}

# PACKAGES 


# dashboard layout, widgets, and uploading

library( flexdashboard )
library( shiny )
library( rsconnect )

# data wrangling 

library( dplyr )


# formatting output

library( DT )
library( pander )
library( knitr )
library( stargazer )


# maps

library( ggmap )
library( leaflet )
library( viridis )
library( geojsonio )
library( sp )
library( sf )
library( tmap )
library( pals )
library( rgdal )
library( mclust )

#other
library(tidycensus)
library(tidyverse)
library(viridis)
library(plyr)
library(gtools)
```



```{r, include=FALSE}

# DATA STEPS

# load dorling cartogram from github
# map already contains census data and groups from clustering 
census_api_key("554b96cd2974f90056633a65d99ece26bcf12905")

github.url <- "https://raw.githubusercontent.com/jmacost5/slc/main/slc_dorling.geojson"
slc <- geojson_read( x=github.url,  what="sp" )

# reproject the map 
slc2 <- spTransform( slc, CRS("+init=epsg:3395") )

# convert the sp map format to 
# an sf (simple features) format:
# ggmap requires the sf format
slc.sf <- st_as_sf( slc2 )

# separate out the data frame from the map
d <- as.data.frame( slc.sf )

```


Community Demographics  
=====================================  

Inputs {.sidebar}
-------------------------------------

```{r}
these.variables <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")

value <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")

dd.name <-c("Percent White not Hispanic", "Percent Black not Hispanic","Percent Hispanic", "Percent Native American", "Percent Forgin Born", "Percent Speaking other Language at home older than 5", "Percent with only Highschool Diploma ", "Percent with College Degree ", "Percent Unemployed",  "Percent Females in Labor Force", "Percent Professional Workers", "Percent Manufactor Workers", "Percent Veterans", "Percent Self Employed", "Median Household Income", "Per capita Income", "Percent in Poverty", "Percent Owner Occupied", 
"Percent Percent Vacancy",  "Percent Multifamily Units", "Median Rent", "Median Home Value",  "Percent structures 30 years old", "Percent 10 years or less", "Percent  under 18 years old", "Percent older than 60", "Percent older than 75", "Percent Married", "Percent Widowed/Separated", "Percent Female Head of Household")

x <- dd.name
names(x) <- value
temp.names <- paste0( dd.name )

radioButtons( inputId="demographics", 
              label = h3("Census Variables"),
              # choices = these.variables, 
              choiceNames=temp.names,
              choiceValues=these.variables,
              selected="pnhwht12")


# Adding interpretable variable names
# from the data dictionary:
# add a name attribute for each variable
# 
# value <- c(1,2,3)
# dd.name <- c("one","two","three")
# 
# x <- dd.name
# names(x) <- value
#
# dd names and values linked
# names( x[2] )
#
# can now get the label using the value
# using the name attributes 
# x[ "two" ]
#
# to add labels to the maps
# use the radio button value 
# to get the data dictionary label: 
#
# x[ input$demographics ]
```



Row {.tabset}
-------------------------------------


### Choropleth Map


```{r}

renderPlot({

# split the selected variable into deciles 
  
get_data <- 
  reactive({
            slc.sf <- 
            slc.sf %>% 
             mutate( q = ntile( get(input$demographics ), 10 ) )
            
          })

  
ggplot( get_data() ) +
    geom_sf( aes( fill = q ), color=NA ) +
    coord_sf( datum=NA ) +
    labs( title = paste0( "Choropleth of Select Demographics: ", toupper(input$demographics) ),
          caption = "Source: Harmonized Census Files",
          fill = "Population Deciles" ) +
    scale_fill_gradientn( colours=rev(brewer.ylorbr(10)), guide = "colourbar" ) + 
    xlim( xmin = -12533000, xmax = -12333000 ) +
ylim( ymin = 4934000, ymax = 5034000 )

})

#xmin = -12300000, xmax = -12600000
#ymax = 5050000, ymin = 4900000
```


### Variable Distribution 



```{r}

renderPlot({

# extract vector x from the data frame 
# x <-  d[ "pnhwht12" ] %>% unlist()

  
get_variable_x <- reactive({ d[ input$demographics ] })

x <- get_variable_x() %>% unlist()

cut.points <- quantile( x, seq( 0, 1, 0.1 ) )

hist( x, breaks=50, 
      col="gray", border="white", yaxt="n",
      main=paste0( "Histogram of variable ", toupper( input$demographics ) ),
      xlab="orange lines represent decile cut points" )

abline( v=cut.points, col="darkorange", lty=3, lwd=2 )


})
```




Neighborhoods  
===================================== 

### Clusters  

```{r}

# define the bounding box corners 
bb <- st_bbox( c( xmin = -12533000, xmax = -12333000 , 
                  ymin = 4934000, ymax = 5034000 ), 
               crs = st_crs("+init=epsg:3395"))
#xmax = -12300000, xmin = -12600000
#ymax = 5050000, ymin = 4900000

# ADD YOUR CUSTOM LABELS TO THE CLUSTERS 

slc2$cluster[ slc2$cluster == "1" ] <- "Salty"
slc2$cluster[ slc2$cluster == "2" ] <- "Swimmers"
slc2$cluster[ slc2$cluster == "3" ] <- "Climbers"

renderTmap({

tmap_mode("view")
tm_basemap( "CartoDB.Positron" )
tm_shape( slc2, bbox=bb ) + 
  tm_polygons( col="cluster", palette="Accent",
               title="Community Types" ) 

})
```




NH Change 2000-2010  
===================================== 



Inputs {.sidebar}
-------------------------------------

```{r}
button.variables <- c("mhv.00","mhv.10","mhv.change","mhv.growth")

button.values <- c("mhv.00","mhv.10","mhv.change","mhv.growth")

button.labels <- c("Median Home Value 2000","Median Home Value 2010","Value Change 2000-2010","Growth in Home Value")


radioButtons( inputId="home.value", 
              label = h3("Home Values"),
              # choices = these.variables, 
              choiceNames=button.labels,
              choiceValues=button.values,
              selected="mhv.00")


```




Row {.tabset}
-------------------------------------



### Median Home Values


```{r}

renderPlot({

# split the selected variable into deciles 
  
get_data <- 
  reactive({
             slc.sf <- 
             slc.sf %>% 
             mutate( q = ntile( get(input$home.value), 10 ) )  
          })

  
ggplot( get_data() ) +
    geom_sf( aes( fill = q ), color=NA ) +
    coord_sf( datum=NA ) +
    labs( title = paste0( "Spatial Distribution of Home Values: ", toupper(input$home.value) ),
          caption = "Source: Harmonized Census Files",
          fill = "Home Value Deciles" ) +
    scale_fill_gradientn( colours=rev(brewer.ylorbr(10)), guide = "colourbar" ) + 
   xlim( xmin = -12533000, xmax = -12333000 ) +
ylim( ymin = 4934000, ymax = 5034000 )

})
#xmin = -12300000, xmax = -12600000, 
#ymax = 5050000, ymin = 4900000

```


### Variable Distribution 

```{r}
renderPlot({

# extract vector x from the data frame 
# x <-  d[ "pnhwht12" ] %>% unlist()

  
get_variable_x <- reactive({ d[ input$home.value ] })

x <- get_variable_x() %>% unlist() %>% as.numeric()

cut.points <- quantile( x, seq( 0, 1, 0.1 ) )

hist( x, breaks=50, 
      col="gray", border="white", yaxt="n",
      main=paste0( "Histogram of ", toupper( input$home.value ) ),
      xlab="orange lines represent decile cut points" )

abline( v=cut.points, col="darkorange", lty=3, lwd=2 )


})
```  







Drivers of Change   
===================================== 



Inputs {.sidebar}
-------------------------------------

```{r}

button.labels <- c("Median Home Value 2000","Median Home Value 2010","Value Change 2000-2010","Growth in Home Value")
button.values <- c("mhv.00","mhv.10","mhv.change","mhv.growth")
  
radioButtons( inputId="dv", 
              label = h3("Select Your Dependent Variable"),
              choiceNames=button.labels,
              choiceValues=button.values,
              selected="mhv.change")


covariates <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")

covariate.labels <-c("Percent White not Hispanic", "Percent Black not Hispanic","Percent Hispanic", "Percent Native American", "Percent Forgin Born", "Percent Speaking other Language at home older than 5", "Percent with only Highschool Diploma ", "Percent with College Degree ", "Percent Unemployed",  "Percent Females in Labor Force", "Percent Professional Workers", "Percent Manufactor Workers", "Percent Veterans", "Percent Self Employed", "Median Household Income", "Per capita Income", "Percent in Poverty", "Percent Owner Occupied", 
"Percent Percent Vacancy",  "Percent Multifamily Units", "Median Rent", "Median Home Value",  "Percent structures 30 years old", "Percent 10 years or less", "Percent  under 18 years old", "Percent older than 60", "Percent older than 75", "Percent Married", "Percent Widowed/Separated", "Percent Female Head of Household") 



checkboxGroupInput( inputId="covariates", 
              label = h3("Select Variables for Your Model"),
              #choices = covariates,
              choiceNames=covariate.labels,
              choiceValues=covariates,
              selected=c("pnhwht12","pprof12","pvac12") )

```




Row {.tabset}
-------------------------------------



### Predicting Change 

```{r, results="asis"}

# RUNNING A REGRESSION WITH USER INPUTS
#
# create a formula object
# by constructing the formula from user selections 
# as a string then casting as a formula object 
# x.s <- c("x1","x2","x3" )
# formula.text <- paste( "y", " ~ ", paste0( x.s, collapse=" + ") )
# formula.object <- as.formula( formula.text )
#
# lm( formula.object, data=d )
#
# make sure all variables are in d


  
# check boxes return vectors 

get_dv <- reactive({ input$dv })
get_covariates <-  reactive({ input$covariates })


renderUI({
  dv <- get_dv()
covariates <- get_covariates() 

formula.text <- paste0( dv, " ~ ", paste( covariates, collapse=" + " ) )
fo <- as.formula( formula.text )

m <- lm( fo, data=d )


# HTML( "<div style='width: 60%; margin: 0px auto;'>" )
HTML( 
  
  c("<br><br><br>",
  "<div type='regression' style='width: 60%; margin: 0px auto;'>",
  stargazer( m, type="html", omit.stat=c("rsq","f") ),
  "</div>",
  "<br><br><br>"
  )
  
  )


})

# HTML( reg.table )
```



### Correlation Plots 



```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex <- 0.8/strwidth(txt)
    
    test <- cor.test(x,y)
    # borrowed from printCoefmat
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE,
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " "))
    
    text(0.5, 0.5, txt, cex = 1.5 )
    text(.7, .8, Signif, cex=cex, col=2)
}

panel.smooth <- function (x, y, col = par("col"), bg = NA, pch = par("pch"), 
    cex = 0.5, col.smooth = "orange", span = 2/3, iter = 3, ...) 
{
    points(x, y, pch = 19, col = gray(0.7,0.2), bg = bg, cex = cex)
    ok <- is.finite(x) & is.finite(y)
    if (any(ok)) 
        lines(stats::lowess(x[ok], y[ok], f = span, iter = iter), 
            col = col.smooth, lwd=2, ...)
}

# custom plot
jplot <- function( x1, x2, lab1="", lab2="", draw.line=T, ... )
{

    plot( x1, x2,
          pch=19, 
          col=gray(0.6, alpha = 0.2), 
          cex=0.5,  
          bty = "n",
          xlab=lab1, 
          ylab=lab2, cex.lab=1.5,
        ... )

    if( draw.line==T ){ 
        ok <- is.finite(x1) & is.finite(x2)
        lines( lowess(x2[ok]~x1[ok]), col="orange", lwd=3 ) }

}
```




```{r}
renderPlot({
  pairs (d[input$covariates], upper.panel= panel.cor, lower.panel=panel.smooth )
})
```





<style>

.chart-shim { overflow: auto; }
 
table{
   border-spacing:1px;
   margin-top:30px;
   margin-bottom:30px;
   margin-left: auto;
   margin-right: auto;
   align:center} 

td{ padding: 6px 10px 6px 10px } 

th{ text-align: left; } 

</style>
